Improving \textsl{Gaia} parallaxes with a data-driven model of the color--magnitude diagram

### 2017-02-21 paperDistance/paper.tex
intro content

Here we deliver a demonstration-of-concept project that shows that a
data-driven model of the \gaia\ data could in principle deliver very
precise distances, without any input of the physics of stellar
evolution, interiors, or photospheres...[DWH CONTENT]

For our data-driven model of the color--magnitude diagram (CMD), we use
an Empirical-Bayes approach known as Extreme Deconvolution (DWH CITE).
This method deconvolves heteroskedastic data to derive an estimate
of the noise-free or low-noise distribution that \emph{would have}
been observed with far better data.
This method has been used in astrophysics
previously to model the velocity distribution of stars in the Solar
Neighborhood (DWH CITE BOTH) and to perform quasar target selection
in Sloan Digital Sky Survey data (DWH CITE BOTH).
Its advantage over other deconvolution methods is that it takes as input,
and handles in a principled way, heteroskedastic data.
Its principle disadvantage for the present purposes is that it requires
or implies a strictly Gaussian noise model.
In the below, this requires us to transform the CMD to a space in which
the noise is approximately Gaussian.

@@

\begin{thebibliography}{}
\end{thebibliography}

>>>>

Here we deliver a demonstration-of-concept project that shows that a
data-driven model of the \gaia\ data could in principle deliver very
precise distances, without any input of the physics of stellar
evolution, interiors, or photospheres.
Most stellar models used to generate photometric distances or photometric
parallaxes have been physical models.
Because there are small issues with physical models of stars, leading to small
color and luminosity differences with the data, these models produce distance
estimates that are biased.
Besides, they build a long list of physical assumptions---assumptions about
how stars work and how to model them numerically---into the distance estimates.
That is, it is impossible (when using physical models)
to deliver distance estimates that involve minimal in assumptions.

However, the use of physical models for distance estimation is not necessary.
Because there are stars at all luminosities with good parallax information,
it is possible to build a stellar photometric model from the data themselves.
Different stars are observed at different levels of parallax precision,
so it requires relatively high technology to build this data-driven model
using all the data available, fairly and responsibly.

Here we build and use just such a data-driven model.
In particular, we use all of the \gaia\ \tgas\ data, matched to the \tmass\ 
photometry, to make a model of the noise-free (or low-noise)
color--magnitude diagram (CMD) of stars.
We use
an Empirical-Bayes approach known as Extreme Deconvolution (DWH CITE).
This method deconvolves heteroskedastic data to derive an estimate
of the noise-free or low-noise distribution that \emph{would have}
been observed with far better data.
This method has been used in astrophysics
previously to model the velocity distribution of stars in the Solar
Neighborhood (DWH CITE BOTH) and to perform quasar target selection
in Sloan Digital Sky Survey data (DWH CITE BOTH).
Its advantage over other deconvolution methods is that it takes as input,
and handles in a principled way, heteroskedastic data.
Its principle disadvantage for the present purposes is that it requires
or implies a strictly Gaussian noise model.
In the below, this requires us to transform the CMD to a space in which
the noise is approximately Gaussian.

@@

%\begin{thebibliography}{}
%\end{thebibliography}
